---
title: "Linear Models 3 (Lab)"
output:
  html_document:
    df_print: paged
---

## Pre-lab

```{r}
# import libraries
library("tidyverse")  # for wrangling, plotting, etc.
library("knitr")      # for knitting RMarkdown 
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 

# load data
data <- read.csv("data_pc_soa.csv") %>%
                  mutate(cue_type = factor(cue_type), soa = factor(soa))

# visualize data at SOA of 100 ms only: 
data_mod <- data[data$soa == 100,]
ggplot(data = data_mod,
       aes(x = cue_type, y = pc)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(height = 0, width = 0.1)) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange",
               size = 1) + 
  stat_summary(fun = "mean",
               geom = "point",
               shape = 21,
               size = 4) +
    labs(y = "Proportion Correct (%)", x = "Trial Type")
```


## One-way ANOVA

Perform an one-way ANOVA on proportion correct by holding SOA constant at **100 ms** using only **cue_type** as the factor (aka. using `data_mod`). Print the summary of the ANOVA results.

```{r}
aov1 <- aov(pc ~ cue_type, data_mod)
summary(aov1)
```

## Linear Model Approach

Fit a compact model and a augmented model to `data_mod`. And plot each model with residuals.

```{r}
# fit models
fit_c = lm(formula = pc ~ 1, data_mod)
fit_a = lm(formula = pc ~ cue_type, data_mod)
```

```{r}
# plot compact
set.seed(1)
data.plot = data_mod %>% 
  mutate(cue_type_jitter = 1 + runif(n(), min = -0.15, max = 0.15))

data.augment = fit_c %>% 
  augment() %>% 
  clean_names() %>% 
  bind_cols(data.plot %>% 
              select(cue_type, cue_type_jitter))

ggplot(data = data.plot, 
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     fill = cue_type)) + 
  geom_hline(yintercept = mean(data$pc)) +
  geom_point(alpha = 0.5) + 
  geom_segment(data = data.augment,
               mapping = aes(xend = cue_type_jitter,
                             yend = fitted),
               alpha = 0.2) +
  labs(y = "Meowmeow beans") + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
# Plot augmented
set.seed(1)
df.plot = data_mod %>% 
  mutate(cue_type_jitter = cue_type %>% 
           as.numeric(),
          cue_type_jitter = cue_type_jitter + runif(n(), min = -0.2, max = 0.2))

df.tidy = fit_a %>% 
  tidy() %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.fns = ~ round(., digits = 2)))

df.augment = fit_a %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% 
              select(cue_type_jitter))

ggplot(data = df.plot,
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     color = cue_type)) + 
  geom_point(alpha = 0.8) +
  geom_segment(data = NULL,
               mapping = aes(x = 0.6,
                             xend = 1.4,
                             y = df.tidy$estimate[1],
                             yend = df.tidy$estimate[1]),
               color = "lightpink",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "lightblue",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[3],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[3]),
               color = "peachpuff",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = cue_type_jitter,
                   y = pc,
                   yend = fitted),
               alpha = 0.3) +
  labs(y = "Proportion Correct (%)") + 
  scale_color_manual(values = c("lightpink", "lightblue", "peachpuff")) + 
  scale_x_continuous(breaks = 1:3, labels = c("invalid", "neutral", "valid")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```
```{r}
# run anova
anova(fit_c, fit_a)
```

## Contrasts

1.  Print the contrast matrix of all independent variables and identify the type of coding (hint: `?contrasts`)

```{r echo=TRUE}
# Contrast matrices for each indp. var
contrasts(data$cue_type) %>% kable()
contrasts(data$soa) %>% kable();
```
  a. Which cue type is the baseline?
  "invalid"
  
2.  Change the contrast matrix of `data$cue_type` so the cue_type is the baseline. (hint: `contrasts(data$cue_type) <- <your contrast matrix>`)

```{r}
  data.new <- data %>% mutate(cue_type = factor(cue_type, levels = c("neutral", "valid", "invalid")))
  contrasts(data.new$cue_type) %>% kable();
```

  a. Redo your linear-model based analysis using the new effect coding:

```{r}
lm.new <- data.new %>% lm(formula = pc ~ cue_type);
lm.new %>% summary() %>% tidy() %>% kable()
```
  
  b. What do you notice about the F-statistic when compared to your original model? Why is this the case?
  
  F-statistic remains the same. Changing the baseline does not affect the test which examines whether there is a differnce between means. 
  
  c. What do you notice about the P-Value of each coefficient?
  
  They have changed. The baseline shifted so the t-statistic that this coefficient is different from 0 will also be different. Each coefficient now represents the effect of either invalid or valid on the baseline neutral trial.
  
  d. Interpret the new coefficients in the context of this new coding
  
  The valid condition increases the performance compared to the neutral condition by about 1.6%. The invalid cue type reduced the proportion of correct trials by about 1.4%.
  
  
## Two-way ANOVA

1.  Visualize the proportion of correct trials (`data$pc`) for both factors (`data$soa` and `data$cue_type`) using boxplots. Let the x-axis represent the two SOAs and use different colors to represent the trial types (hint: `aes(x = soa, y = pc, fill = cue_type)`).

```{r}
# visualize data
data %>%
  ggplot(mapping = aes(x = soa,y = pc, fill = cue_type)) +
  geom_boxplot(notch = FALSE) +
  ylab("Proportion Correct") +
  xlab("SOA (ms)") +
  scale_size_area("Cue Type")
```

2. Run an ANOVA to analyze the effect of both independent variables on the proportion correct

  a. Without interaction
```{r}
aov(data = data, formula = pc ~ cue_type + soa) %>% summary();
```
  b. With interaction
```{r}
aov(data = data, formula = pc ~ cue_type*soa) %>% summary();
```
  
  c. Choose a model that best characterizes the effect

  Both models do not significantly reduce the variance compared to the compact model. However, it seems that the second model does slightly better when the interaction term is added.
  
  d. Report the statistics
  
  We fail to reject the null hypothesis that the mean value of proportion correct is significantly different for different cue types $p = 0.477$. Furthermore, the mean proportion correct is not significantly different for SOAs $F=0.625, p = 0.434, Df = 1$. Finally, the interaction between SOA and cue type is not significant ($F=1.150, p = 0.326$)
  
  