---
title: "Linear Models 3 (Lab)"
output:
  html_document:
    df_print: paged
---

## Introduction

This is a dataset of a classical Posner task where subjects were cued to shift their attention towards one location on the screen. They were then asked to discriminate the orientation of a tilted bar on either the same location they were cued to (valid) or the opposite location from the cued location (invalid). Some trials did not contain a cue (neutral).

## Pre-lab

```{r}
# import libraries
library("tidyverse")  # for wrangling, plotting, etc.
library("knitr")      # for knitting RMarkdown 
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 

# load data
data <- read.csv("data_pc_soa.csv") %>%
                  mutate(cue_type = factor(cue_type), soa = factor(soa))
```
Visualize data at SOA of 100 ms only. x-axis: invalid, neutral, valid; y-axis: pc (proportion correct) with error bars of confidence intervals or standard deviations.
```{r}
data_mod <- data[data$soa == 100,] # extract data of 100 ms SOA 
# complete the code here

```

## One-way ANOVA

Perform an one-way ANOVA on proportion correct by holding SOA constant at **100 ms** using only **cue_type** as the factor (aka. using `data_mod`). Print the summary of the ANOVA results. Use function `aov()`.

```{r}
# perform one-way ANOVA
```

## Linear Model Approach

Fit a compact model and a augmented model to `data_mod`. Use function `lm()`.
```{r}
# fit models
```

Run an ANOVA test on the two models. Use function `anova()`
```{r}
# run ANOVA
```

## Contrasts

1.  Print the contrast matrix of all independent variables and identify the type of coding (hint: `?contrasts`)

```{r echo=TRUE}
# Contrast matrices for each indp. var
```

    a.  Which cue type is the baseline?

2.  Change the contrast matrix of `data$cue_type` so the cue_type is the baseline. (hint: `contrasts(data$cue_type) <- <your contrast matrix>`)

    a.  Redo your linear-model based analysis using the new effect coding:

    b.  What do you notice about the F-statistic when compared to your original model? Why is this the case?

    c.  What do you notice about the P-Value of each coefficient?

    d.  Interpret the new coefficients in the context of this new coding

## Two-way ANOVA

1.  Visualize the proportion of correct trials (`data$pc`) for both factors (`data$soa` and `data$cue_type`) using boxplots. Let the x-axis represent the two SOAs and use different colors to represent the trial types (hint: `aes(x = soa, y = pc, fill = cue_type)`).

```{r}
# visualize data
```

2.  Run an ANOVA to analyze the effect of both independent variables on the proportion correct

    a.  Without interaction

    b.  With interaction

    c.  Choose a model that best characterizes the effect

    d.  Report the statistics
