---
title: "Linear Models 3 (Lab)"
output:
  html_document:
    df_print: paged
---
## Pre-lab
```{r}
# import libraries
library("tidyverse")  # for wrangling, plotting, etc.
library("knitr")      # for knitting RMarkdown 
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 

# load data
data <- read.csv("data_pc_soa.csv") %>%
                  mutate(cue_type = factor(cue_type), soa = factor(soa))

# visualize data at SOA of 100 ms only
data_mod <- data[data$soa == 100,]
ggplot(data = data_mod,
       aes(x = cue_type, y = pc)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(height = 0, width = 0.1)) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange",
               size = 1) + 
  stat_summary(fun = "mean",
               geom = "point",
               shape = 21,
               size = 4) +
  labs(y = "Proportion Correct (%)", x = "Trial Type")
```
## One-way ANOVA
Perform an one-way ANOVA on proportion correct by holding SOA constant at **100 ms** using only **cue_type** as the factor (aka. using `data_mod`). Print the summary of the ANOVA results.
```{r}
aov1 <- aov(pc ~ cue_type, data_mod)
summary(aov1)
```
## Linear Model Approach
Fit a compact model and a augmented model to `data_mod`. And plot each model with residuals.
```{r}
# fit models
fit_c = lm(formula = pc ~ 1, data)
fit_a = lm(formula = pc ~ cue_type, data)
```

```{r}
# plot compact
set.seed(1)
data.plot = data %>% 
  mutate(cue_type_jitter = 1 + runif(n(), min = -0.15, max = 0.15))

data.augment = fit_c %>% 
  augment() %>% 
  clean_names() %>% 
  bind_cols(data.plot %>% 
              select(cue_type, cue_type_jitter))

ggplot(data = data.plot, 
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     fill = cue_type)) + 
  geom_hline(yintercept = mean(data$pc)) +
  geom_point(alpha = 0.5) + 
  geom_segment(data = data.augment,
               mapping = aes(xend = cue_type_jitter,
                             yend = fitted),
               alpha = 0.2) +
  labs(y = "Meowmeow beans") + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
# Plot augmented
set.seed(1)
df.plot = data %>% 
  mutate(cue_type_jitter = cue_type %>% 
           as.numeric(),
          cue_type_jitter = cue_type_jitter + runif(n(), min = -0.2, max = 0.2))

df.tidy = fit_a %>% 
  tidy() %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.fns = ~ round(., digits = 2)))

df.augment = fit_a %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% 
              select(cue_type_jitter))

ggplot(data = df.plot,
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     color = cue_type)) + 
  geom_point(alpha = 0.8) +
  geom_segment(data = NULL,
               mapping = aes(x = 0.6,
                             xend = 1.4,
                             y = df.tidy$estimate[1],
                             yend = df.tidy$estimate[1]),
               color = "lightpink",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "lightblue",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[3],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[3]),
               color = "peachpuff",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = cue_type_jitter,
                   y = pc,
                   yend = fitted),
               alpha = 0.3) +
  labs(y = "Proportion Correct (%)") + 
  scale_color_manual(values = c("lightpink", "lightblue", "peachpuff")) + 
  scale_x_continuous(breaks = 1:3, labels = c("invalid", "neutral", "valid")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

## Two-way ANOVA
Visualize all data in one plot. Both factors should be included (SOA and cue type). 
```{r}
# visualize data
data %>%
  ggplot(mapping = aes(x = pc,y = soa, fill = cue_type)) +
  geom_boxplot(notch = FALSE) +
  coord_flip() +
  xlab("Proportion Correct") +
  ylab("SOA (ms)") +
  scale_size_area("Cue Type")
```