---
title: "Linear Models 3 (Lecture)"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library("knitr")      # for knitting RMarkdown 
library("janitor")    # for cleaning column names
library("tidyverse")  # for wrangling, plotting, etc.
library("broom")      # for tidying up linear models 
theme_set(theme_bw() + #set the theme 
            theme(text = element_text(size = 16))) #set the default text size
```

## Generate dataset

```{r}
#population distribution
set.seed(124)
mu_cat = 50
sig_cat = 6
cat = rnorm(5,mu_cat,sig_cat)

mu_dog = 20
sig_dog = 6
dog = rnorm(5,mu_dog,sig_dog)

mu_hampster = 20
sig_hampster = 6
hampster = rnorm(5,mu_hampster,sig_hampster)

```

## Visualize dataset

```{r}
df<-tibble(cat, dog, hampster)
df<-pivot_longer(df, everything(), names_to = "pet", values_to = "beans") %>% mutate(pet = factor(pet))
ggplot(data = df,
       aes(x = pet, y = beans)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(height = 0, width = 0.1)) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange",
               size = 1) + 
  stat_summary(fun = "mean",
               geom = "point",
               shape = 21,
               size = 4) +
  labs(y = "Meowmeow beans")
```

## One-way ANOVA

```{r}
aov1<-aov(beans~factor(pet), df)
summary(aov1)
```

## Model fitting

```{r}
fit_c = lm(formula = beans ~ 1, data = df)
fit_a = lm(formula = beans ~ pet, data = df)
summary(fit_c)
summary(fit_a)

# plot compact
set.seed(1)
df.plot = df %>% 
  mutate(pet_jitter = 1 + runif(n(), min = -0.15, max = 0.15))

df.augment = fit_c %>% 
  augment() %>% 
  clean_names() %>% 
  bind_cols(df.plot %>% 
              select(pet, pet_jitter))

ggplot(data = df.plot, 
       mapping = aes(x = pet_jitter,
                     y = beans,
                     fill = pet)) + 
  geom_hline(yintercept = mean(df$beans)) +
  geom_point() + 
  geom_segment(data = df.augment,
               mapping = aes(xend = pet_jitter,
                             yend = fitted),
               alpha = 0.8) +
  labs(y = "Meowmeow beans") + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
# Plot augmented
set.seed(1)
df.plot = df %>% 
  mutate(pet_jitter = pet %>% 
           factor() %>%
           as.numeric(),
         pet_jitter = pet_jitter + runif(n(), min = -0.2, max = 0.2))

df.tidy = fit_a %>% 
  tidy() %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.fns = ~ round(., digits = 2)))

df.augment = fit_a %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% 
              select(pet_jitter))

ggplot(data = df.plot,
       mapping = aes(x = pet_jitter,
                     y = beans,
                     color = pet)) + 
  geom_point() +
  geom_segment(data = NULL,
               mapping = aes(x = 0.6,
                             xend = 1.4,
                             y = df.tidy$estimate[1],
                             yend = df.tidy$estimate[1]),
               color = "grey",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "grey",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[3],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[3]),
               color = "grey",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = pet_jitter,
                   y = beans,
                   yend = fitted),) +
  labs(y = "Meowmeow beans") + 
  scale_color_manual(values = c("grey", "grey", "grey")) + 
  scale_x_continuous(breaks = 1:3, labels = c("cat", "dog", "hampster")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

## Performance ANOVA on two models
```{r}
print(anova(fit_c, fit_a))
```

# Plot between group variance
```{r}

ggplot(data = df.plot, 
       mapping = aes(x = pet_jitter,
                     y = beans,
                     fill = pet)) + 
  geom_hline(yintercept = mean(df$beans)) +
  geom_segment(data = NULL,
               mapping = aes(x = 0.6,
                             xend = 1.4,
                             y = df.tidy$estimate[1],
                             yend = df.tidy$estimate[1]),
               color = "lightpink",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "lightblue",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[3],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[3]),
               color = "lightgreen",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = pet_jitter,
                   y = mean(df$beans),
                   yend = mean(df$beans)),) +
  labs(y = "Meowmeow beans") + 
  scale_color_manual(values = c("lightpink", "lightblue", "lightgreen")) + 
  scale_x_continuous(breaks = 1:3, labels = c("cat", "dog", "hampster")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

### Dummy Coding (Default for `factors` in R)
```{r}
# Print the contrast matrix
df.dum <- df %>% filter(pet %in% c("dog", "cat")) %>% mutate(pet = factor(pet))
lm.dum <- df.dum %>% lm(formula = beans ~ pet);

df.dum %>%
  pull(pet) %>% contrasts() %>%
  kable()
```
### Sum Coding

1.  Extract just Dog and Cat

```{r}
df.sum_coded <- df %>% filter(pet %in% c("dog", "cat")) %>%
  mutate(pet = factor(pet))
```

2.  Set the coding (contrast) to `contr.sum`

```{r}
contrasts(df.sum_coded$pet) <- contr.sum(2) # n-levels = 2
```

3.  (Optional) Change column names so we can easily interpret later

```{r}
colnames(attr(df.sum_coded$pet, "contrasts")) <- c("->vs_GrandMean")
contrasts(df.sum_coded$pet) %>% kable()
```

4.  Run the analysis

```{r}
lm.sum <- df.sum_coded %>% lm(formula = beans ~ pet);
  
lm.sum %>%
  summary() %>%
  tidy() %>%
  kable()
```

## Are F-statistics the same?

### Dummy coded
```{r}
lm.dum %>% anova() %>% tidy() %>% kable()
```

### Sum coded
```{r}
lm.sum %>% anova() %>% tidy() %>% kable()
```

# How do the models differ?
## Predictions based on the model
```{r}
pred.dum <- predict(lm.dum, data.frame(pet = levels(df.dum$pet))) %>% round(2);
pred.sum <- predict(lm.sum, data.frame(pet = levels(df.sum_coded$pet))) %>% round(2);
```

## Dummy Coding Visualization
```{r}
ggplot(mapping = aes(x = factor(c(0, 1)), y = pred.dum, label = pred.dum, vjust=1, hjust=-1)) +
  geom_point() + geom_text() +
  geom_hline( mapping = aes(yintercept = pred.dum[1]), color = "red") + 
  geom_text(mapping = aes(0, pred.dum[1], label = paste("B_0 = ", pred.dum[1] , " (Cat Mean)"), vjust= 1.3, hjust = -0.1)) +
  geom_segment(x = factor(1), y = pred.dum[1], xend = factor(1), yend = pred.dum[2], size = 1, arrow = arrow()) + 
  labs(title = "Dummy Coded", x = "Value Plugged into Model", y = "Prediction (Beans)");
```

## Sum Coding Visualization
```{r}
mu.dogCat <- round(mean(df.sum_coded$beans), 2)
ggplot(mapping = aes(x = factor(c("-1", "1")), y = pred.dum, label = pred.sum, vjust=1, hjust=-1)) +
  geom_point() + geom_text() +
  geom_hline( mapping = aes(yintercept = mu.dogCat), color = "green") + 
  geom_text(mapping = aes(0, mu.dogCat, label = paste("B_0 = ", mu.dogCat , " (Grand Mean)"), vjust= -1, hjust = -0.1)) +
  geom_segment(x = "-1", y = mu.dogCat, xend = "-1", yend = pred.sum[1], size = 1, arrow = arrow()) + 
  geom_segment(x = factor(1), y = mu.dogCat, xend = factor(1), yend = pred.sum[2], size = 1, arrow = arrow()) + 
  labs(title = "Sum Coded", x = "Value Plugged into Model", y = "Prediction (Beans)");
```
## Comapring the Coefficients
```{r, echo=FALSE}
lm.dum %>% summary() %>% tidy() %>% kable
lm.sum %>% summary() %>% tidy() %>% kable
```

# Categorical Variable with 3 Levels
## Dummy Coding
### Contrasts
```{r, echo=FALSE}
contrasts(df$pet) %>% kable();
```

### Coefficients
```{r, echo = FALSE}
lm.dum <- df %>% lm(formula = beans ~ pet);
summary(lm.dum) %>% tidy() %>% kable();
```
### Model
$$ \text{Beans} = `r lm.dum$coefficients[1]` `r lm.dum$coefficients[2]` (\text{Dog}) `r lm.dum$coefficients[3]` (\text{Hamster}) $$

### Visualize

```{r, echo=FALSE}
ggplot(mapping = aes(x = df$pet, y = predict(lm.dum))) +
  geom_point() + 
  geom_hline(yintercept = lm.dum$coefficients[1], color = "green") + 
  geom_text(label = round(predict(lm.dum), 2), vjust = 1) +
  geom_segment(aes(x = df$pet[2], y = predict(lm.dum)[1], xend = df$pet[2], yend = predict(lm.dum)[2]), arrow = arrow()) +
  geom_segment(aes(x = df$pet[3], y = predict(lm.dum)[1], xend = df$pet[3], yend = predict(lm.dum)[3]), arrow = arrow()) +
  xlab("Pet") +
  ylab("Prediction") +
  ggtitle("Dummy Coding");
```

## Sum Coding

### Contrasts
```{r, echo=FALSE}
df.sum <- df %>% mutate(pet = factor(pet));
contrast.sum <- contr.sum(3); 
colnames(contrast.sum) <- c(":cat_vs_mean", ":dog_vs_mean");
rownames(contrast.sum) <- levels(df$pet);
contrast.sum %>% kable();
```

### Coefficients
```{r, echo=FALSE}
lm.sum <- df.sum %>% lm(formula = beans ~ pet, contrasts = list(pet = contrast.sum));
summary(lm.sum) %>% tidy() %>% kable();
```

### Model
$$ \text{Beans} = `r lm.sum$coefficients[1]` + `r lm.sum$coefficients[2]` X_1 `r lm.sum$coefficients[3]` X_2 $$

### Visualization
```{r, echo=FALSE}
ggplot(mapping = aes(x = df$pet, y = predict(lm.sum))) +
  geom_point() + 
  geom_hline(yintercept = lm.sum$coefficients[1], color = "red") + 
  geom_text(label = round(predict(lm.sum), 2), vjust = 1) +
  geom_segment(aes(x = df$pet[1], y = mean(df$beans), xend = df$pet[1], yend = predict(lm.sum)[1]), arrow = arrow()) +
  geom_segment(aes(x = df$pet[2], y = mean(df$beans), xend = df$pet[2], yend = predict(lm.sum)[2]), arrow = arrow()) +
  geom_segment(aes(x = df$pet[3], y = mean(df$beans), xend = df$pet[3], yend = predict(lm.sum)[3]), arrow = arrow()) +
  xlab("Pet") +
  ylab("Prediction") +
  ggtitle("Sum Coding");
```

# Extending beyond 3 levels

## Dummy

```{r, results='asis'}
pets <- c("cat", "dog", "hamster", "turtle", "rabbit", "fish");
for (x in 4:6) {
  cat(paste("### ", x, " variables: "));
  temp <- contr.treatment(x);
  colnames(temp) <- paste("x", as.character(1:(x-1)), sep = '');
  rownames(temp) <- as.character(pets[1:x]);
  print(temp %>% kable(row.names = TRUE));
}
```
## Sum Coding

```{r, results='asis'}
for (x in 4:6) {
  cat(paste("### ", x, " variables: "));
  temp <- contr.sum(x);
  colnames(temp) <- paste("x", as.character(1:(x-1)), sep = '');
  rownames(temp) <- as.character(pets[1:x]);
  print(temp %>% kable(row.names = TRUE));
}
```

# Other types of factor coding

## Helmert Coding
```{r, results='asis'}
for (x in 4:6) {
  cat(paste("### ", x, " variables: "));
  temp <- contr.helmert(x);
  colnames(temp) <- paste("x", as.character(1:(x-1)), sep = '');
  rownames(temp) <- as.character(1:x);
  print(temp %>% kable(row.names = TRUE));
}
```



