---
title: "Linear Models 3 (Lab)"
output:
  html_document:
    df_print: paged
---

## Pre-lab

```{r}
# import libraries
library("tidyverse")  # for wrangling, plotting, etc.
library("knitr")      # for knitting RMarkdown 
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 

# load data
data <- read.csv("data_pc_soa.csv") %>%
                  mutate(cue_type = factor(cue_type), soa = factor(soa))

# visualize data at SOA of 100 ms only: 
data_mod <- data[data$soa == 100,]
ggplot(data = data_mod,
       aes(x = cue_type, y = pc)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(height = 0, width = 0.1)) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange",
               size = 1) + 
  stat_summary(fun = "mean",
               geom = "point",
               shape = 21,
               size = 4) +
    labs(y = "Proportion Correct (%)", x = "Trial Type")
```


## One-way ANOVA

Perform an one-way ANOVA on proportion correct by holding SOA constant at **100 ms** using only **cue_type** as the factor (aka. using `data_mod`). Print the summary of the ANOVA results.

```{r}
aov1 <- aov(pc ~ cue_type, data_mod)
summary(aov1)
```

## Linear Model Approach

Fit a compact model and a augmented model to `data_mod`. And plot each model with residuals.

```{r}
# fit models
fit_c = lm(formula = pc ~ 1, data)
fit_a = lm(formula = pc ~ cue_type, data)
```

```{r}
# plot compact
set.seed(1)
data.plot = data %>% 
  mutate(cue_type_jitter = 1 + runif(n(), min = -0.15, max = 0.15))

data.augment = fit_c %>% 
  augment() %>% 
  clean_names() %>% 
  bind_cols(data.plot %>% 
              select(cue_type, cue_type_jitter))

ggplot(data = data.plot, 
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     fill = cue_type)) + 
  geom_hline(yintercept = mean(data$pc)) +
  geom_point(alpha = 0.5) + 
  geom_segment(data = data.augment,
               mapping = aes(xend = cue_type_jitter,
                             yend = fitted),
               alpha = 0.2) +
  labs(y = "Meowmeow beans") + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
# Plot augmented
set.seed(1)
df.plot = data %>% 
  mutate(cue_type_jitter = cue_type %>% 
           as.numeric(),
          cue_type_jitter = cue_type_jitter + runif(n(), min = -0.2, max = 0.2))

df.tidy = fit_a %>% 
  tidy() %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.fns = ~ round(., digits = 2)))

df.augment = fit_a %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% 
              select(cue_type_jitter))

ggplot(data = df.plot,
       mapping = aes(x = cue_type_jitter,
                     y = pc,
                     color = cue_type)) + 
  geom_point(alpha = 0.8) +
  geom_segment(data = NULL,
               mapping = aes(x = 0.6,
                             xend = 1.4,
                             y = df.tidy$estimate[1],
                             yend = df.tidy$estimate[1]),
               color = "lightpink",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "lightblue",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[3],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[3]),
               color = "peachpuff",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = cue_type_jitter,
                   y = pc,
                   yend = fitted),
               alpha = 0.3) +
  labs(y = "Proportion Correct (%)") + 
  scale_color_manual(values = c("lightpink", "lightblue", "peachpuff")) + 
  scale_x_continuous(breaks = 1:3, labels = c("invalid", "neutral", "valid")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

## Contrasts

1.  Print the contrast matrix of all independent variables and identify the type of coding (hint: `?contrasts`)

```{r echo=TRUE}
# Contrast matrices for each indp. var
contrasts(data$cue_type)
contrasts(data$soa)
```
  a. Which cue type is the baseline?
  
2.  Change the contrast matrix of `data$cue_type` so the cue_type is the baseline. (hint: `contrasts(data$cue_type) <- <your contrast matrix>`)

  a. Redo your linear-model based analysis using the new effect coding:
  
  b. What do you notice about the F-statistic when compared to your original model? Why is this the case?
  
  c. What do you notice about the P-Value of each coefficient?
  
  d. Interpret the new coefficients in the context of this new coding
  
## Two-way ANOVA

1.  Visualize the proportion of correct trials (`data$pc`) for both factors (`data$soa` and `data$cue_type`) using boxplots. Let the x-axis represent the two SOAs and use different colors to represent the trial types (hint: `aes(x = soa, y = pc, fill = cue_type)`).

```{r}
# visualize data
data %>%
  ggplot(mapping = aes(x = soa,y = pc, fill = cue_type)) +
  geom_boxplot(notch = FALSE) +
  ylab("Proportion Correct") +
  xlab("SOA (ms)") +
  scale_size_area("Cue Type")
```

2. Run an ANOVA to analyze the effect of both independent variables on the proportion correct

  a. Without interaction
  
  b. With interaction
  
  c. Choose a model that best characterizes the effect
  
  d. Report the statistics