---
title: "Linear Models I - Week 2"
author: "Yue Zhang and Oviya Mohan"
date: "8/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Load Req. Libraries 
```{r}
library("knitr")      # for knitting RMarkdown 
#library("kableExtra") # for making nice tables
#library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("tidyverse")  # for wrangling, plotting, etc. 
```

## Modelling Data - could be the pre-lab component 

### Simple example with Augmented and Compact Model

## Linear Regression

### Create Sample 
```{r}
# make example reproducible 
set.seed(1)

# set the sample size
n_samples = 10

# generate correlated data
df.regression = tibble(chocolate = runif(n_samples, min = 0, max = 100),
                       happiness = chocolate * 0.5 + rnorm(n_samples, sd = 15))

# plot the data 
ggplot(data = df.regression,
       aes(x = chocolate,
           y = happiness)) +
  geom_point(size = 3)
```

### Define and fit the compact model (Model C) 

### Define and fit the augmented model (Model A) 

### Caluclate Sum of Sqaured errors of each model and visulize it 

### Calculate the F statistic to determine whether PRE (Proportion Reduction in Error) is significant 

## Example with Real Data Set 

## Correlation (text book has correlation before regression - could skip in lab, or simply the problem, don't need this level of detail...) 
```{r}
# make example reproducible 
set.seed(1)

n_samples = 20

# create correlated data
df.correlation = tibble(x = runif(n_samples, min = 0, max = 100),
                        y = x + rnorm(n_samples, sd = 15))

# plot the data
ggplot(data = df.correlation,
       mapping = aes(x = x,
                     y = y)) + 
  geom_point(size = 2) +
  labs(x = "chocolate",
       y = "happiness")
```

### Variance 

### Covariance 

Why are they generating a different data set for each of these calculations - does it not make sense to have one data set and calculate all these values for it? 

### Spearman's rank order correlation 

#### Show that Spearman’s ρ is equivalent to Pearson’s r applied to ranked data.
