---
title: "Linear Models I - Week 2"
author: "Yue Zhang and Oviya Mohan"
date: "8/25/2022"
output:
  html_document: default
  pdf_document: default
---

## Introduction

This dataset looks at the effect of saccade landing errors on pre-saccadic attention. Saccades are rapid eye movements that happen in between fixations. Pre-saccadic attention is a shift in spatial attention from fixation to the future saccade goal prior to the onset of the saccade. Shift in attention is measured by the performance of detecting or discrimating a visual target at the saccade goal location. Here, we are investigating how the precision of saccade landing affect the performance by looking at how the distance between the actual saccade landing and the visual target changes the reaction time of discriminating the visual target. 

## Pre-lab
```{r,warning=FALSE,message=FALSE}
# install and attach libraries
# install.packages("janitor")
library("knitr")      # for knitting RMarkdown 
#library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("tidyverse")  # for wrangling, plotting, etc. 
```

```{r}
#load data
dataframe <- read.csv("msData.csv") 
#plot data
ggplot(data = dataframe,
       aes(x = dist,
           y = rt)) +
  geom_point(size = 1)
```

## Step1: Hypotheses
$H_{0}$ =

$H_{1}$ =

## Step2: Fit Models

### Define and fit the compact model (Model C):

$$Y_{i} = \beta_{0} + \epsilon_{i}$$

```{r}
# fit the compact model
lm.compact = lm(rt ~ 1, data = dataframe)

# store the results of the model fit in a data frame
df.compact = tidy(lm.compact)

# plot the data with model prediction
ggplot(data = dataframe,
       aes(x = dist,
           y = rt)) +
  geom_hline(yintercept = df.compact$estimate,
             color = "red",
              size = 1) +
  geom_point(size = 1, shape = 5) 
```

### Define and fit the augmented model (Model A):

$$Y_{i} = \beta_{0} + \beta_{1}X_{i}+ \epsilon_{i}$$

```{r}
# fit the augmented model
lm.augmented = lm(rt ~ dist, data = dataframe)

# store the results of the model fit in a data frame
df.augmented = tidy(lm.augmented)

# plot the data with model prediction
ggplot(data = dataframe,
       aes(x = dist,
           y = rt)) +
  geom_abline(intercept = df.augmented$estimate[1],
              slope = df.augmented$estimate[2],
              color = "blue",
              size = 1) +
  geom_point(size = 1, shape = 5) 
```

## Step3: Compare Models

### Caluclate residuals and SSE of Model C
```{r}
# create a data frame that contains the residuals 
df.compact_model = augment(lm.compact) %>% 
  clean_names() %>% 
  left_join(dataframe, by = "rt")

# plot model prediction with residuals
ggplot(data = df.compact_model,
       aes(x = dist,
           y = rt)) +
  geom_hline(yintercept = df.compact$estimate,
             color = "red",
              size = 1) +
  geom_segment(aes(xend = dist,
                   yend = df.compact$estimate),
               color = "red") + 
  geom_point(size = 1, shape = 5) 

# calculate the sum of squared errors
df.compact_model %>% 
  summarize(SSE = sum(resid^2))
```

### Caluclate residuals and SSE of Model A
```{r}

# create a data frame that contains the residuals 
df.augmented_model = augment(lm.augmented) %>% 
  clean_names() %>% 
  left_join(dataframe, by = c("rt", "dist"))

# plot model prediction with residuals
ggplot(data = df.augmented_model,
       aes(x = dist,
           y = rt)) +
  geom_abline(intercept = df.augmented$estimate[1],
              slope = df.augmented$estimate[2],
             color = "blue",
              size = 1) +
  geom_segment(aes(xend = dist,
                   yend = fitted),
               color = "blue") + 
  geom_point(size = 1, shape = 5) 

# calculate the sum of squared errors
df.augmented_model %>% 
  summarize(SSE = sum(resid^2))
```

### Calculate the F statistic to determine whether PRE (Proportion Reduction in Error) is significant
```{r}
pc = 1 # number of parameters in the compact model  
pa = 2 # number of parameters in the augmented model  
n = 1051 # number of observations

# SSE of the compact model 
sse_compact = df.compact_model %>% 
  summarize(SSE = sum(resid^2))

# SSE of the augmented model
sse_augmented = df.augmented_model %>% 
  summarize(SSE = sum(resid^2))

# Proportional reduction of error 
pre = as.numeric(1 - (sse_augmented/sse_compact))

# F-statistic 
f = (pre/(pa-pc))/((1-pre)/(n-pa))

# p-value
p_value = 1-pf(f, df1 = pa-pc, df2 = n-pa)
print(p_value)

# plot f stats on f distribution
ggplot(data = tibble(x = c(0, 10)),
       mapping = aes(x = x)) +
  stat_function(fun = "df",
                args = list(df1 = pa-pc,
                            df2 = n-pa),
                size = 1) +
  geom_vline(xintercept = f,
             color = "red",
             size = 1)
```

## Step4: Reporting Results
```{r}
summary(lm.augmented)
```
Reporting statistical test results:
Conclusion: 


