---
title: "Linear Models I - Week 2"
author: "Yue Zhang and Oviya Mohan"
date: "8/25/2022"
output:
  html_document: default
  pdf_document: default
---

## Introduction

This dataset looks at the effect of saccade landing errors on pre-saccadic attention. Saccades are rapid eye movements that happen in between fixations. Pre-saccadic attention is a shift in spatial attention from fixation to the future saccade goal prior to the onset of the saccade. Shift in attention is measured by the performance of detecting or discrimating a visual target at the saccade goal location. Here, we are investigating how the precision of saccade landing affect the performance by looking at how the distance between the actual saccade landing and the visual target changes the reaction time of discriminating the visual target. 

## Pre-lab (run the following two chunks before the lab session on 9/14 - post any questions on slack)
```{r,warning=FALSE,message=FALSE}
# install and attach libraries
# install.packages("janitor")
library("knitr")      # for knitting RMarkdown 
#library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("tidyverse")  # for wrangling, plotting, etc. 
```

```{r}
#load data
dataframe <- read.csv("msData.csv") 
#plot data
ggplot(data = dataframe,
       aes(x = dist,
           y = rt)) +
  geom_point(size = 1)
```

## Step1: Hypotheses
$H_{0}$ =

$H_{1}$ =

## Step2: Fit Models

### Define and fit the compact model (Model C):

$$Y_{i} = \beta_{0} + \epsilon_{i}$$

```{r}

```

### Define and fit the augmented model (Model A):

$$Y_{i} = \beta_{0} + \beta_{1}X_{i}+ \epsilon_{i}$$

```{r}

```

## Step3: Compare Models

### Caluclate residuals and SSE of Model C
```{r}

```

### Caluclate residuals and SSE of Model A
```{r}

```

### Calculate the F statistic to determine whether PRE (Proportion Reduction in Error) is significant
```{r}


```

## Step4: Reporting Results
```{r}

```
Reporting statistical test results:

Conclusion: 


